---
title: "R Notebook"
output: html_notebook
---

## Data Setup and Exploration

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

### Get the appropriate libraries
```{r libraries}
library(tidyverse)
library(e1071)
library(mice)
library(Amelia)
library(scales)
library(class)
```

### Load the Data
```{r loadData}
# Get Data from ClassExamples Directory
yeast <- read.csv(file='C:/Users/andre/Documents/GitHub/CST-425/Topic-3/2-1-2022/yeast.data', sep=" ")
col_odd <- seq_len(ncol(yeast)) %% 2
yeast <- yeast[, col_odd == 1]
glimpse(yeast)
```


### Explore the data
```{r glimpse}
# Initial data exploration
glimpse(yeast)
```

```{r explore}
# Noticing the data, we can remove columns 5 and 6 because they have all the same values
yeast %>% select(-5,-6)

yeast <- na.omit(yeast)

# This output should be zero
sum(is.na(yeast))

# Glimpse the data again
glimpse(yeast)
```

### Decide which variable you want to predict and which ones will be the predictors create a subset of the data to include only relevant variables
```{r decideVariables}
```

### Clean data


### Normalize
```{r}

```

### Split the dataset into training and testing subsets
```{r split}
# Set a seed to randomize model generation
set.seed(1234)

# Create a sample of the data. The sample set will just include the body text of the emails, starting at the first element and going to the 80% mark. This will allow the training set sample to be established. A training and testing set ratio should be around 70/30 to 80/20.
sample_set <- sample(1:dim(yeast)[1], dim(yeast)[1]*0.80)

# Split training and testing set using 80:20 split. The training set will use the dimensions from the sample set sample() while the test set will use the inverse of it by using the (-sample_set).
yeast_train <- yeast[sample_set,]
yeast_test <- yeast[-sample_set,]
```

### Use the knn function in R or Python
```{r knn}
# Determine a value for k
kVal <- 10

# Create model
knn(yeast_train, yeast_test, yeast_train$MIT, k = kVal, l = 0, prob = FALSE, use.all = TRUE)
```

### Assess how good is your model