---
title: "Email Spam Filtering"
output: html_notebook
---

## Step 1: Data Setup and Exploration

### Setup Knitr
```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

### Get the appropriate libraries
```{r libraries}
library(tidyverse)
library(e1071)
library(mice)
library(Amelia)
library(scales)

wd <- getwd()
wd
```



### Load the Data
```{r loadData}
# Get Data from ClassExamples Directory
spam <- read_csv(file='EmailSpamFiltering/completeSpamAssassin.csv')
head(spam)
```

# Explore the model by getting a glimpse of the data.
```{r exploreModel}
glimpse(spam)
```

## Step 2: Preparing training and testing sets
```{r model}
# Set a seed to randomize model generation
set.seed(1234)

# Create a sample of the data
sample_set <- sample(1:dim(spam)[1], dim(spam)[1]*0.80)

# Split training and testing set using 75:25 split
spam_train <- spam[sample_set,]
spam_test <- spam[-sample_set,]
```

## Step 3: Build the Bayesian Model (Naive)
```{r buildModel}
# Build the naive model based on training set
spam_mod <- naiveBayes(Label ~ ., data = spam_train, laplace = 1)
spam_mod
```


## Step 4: Improve the Bayesian Model
```{r improveModel}
#Get rid of missing values
#spam.df <- data.frame(spam)
glimpse(spam)

spam[spam == "empty"] <- NA
sum(is.na(spam))
spam.removedNA<- na.omit(spam)


glimpse(spam.removedNA)



#Make all words lowercase to prevent multiple matching
# Lowercase the model
#spam$Body <- tolower(spam$Body)
#spam$Body <- toupper(spam$Body)
head(spam,1)



##Changing Testing and Training set ratios

# Create a sample of the data
sample_set <- sample(1:dim(spam)[1], dim(spam)[1]*0.75)

# Split training and testing set using 75:25 split
spam_train <- spam[sample_set,]
spam_test <- spam[-sample_set,]




```


## Step 5: Interpret the Model Output
```{r interpretModel}
# TK
```


## Step 6: Estimate Errors
```{r errorEstimation}

```


## Step 7: Make predictions using the test set
```{r predict}
spam_pred <- predict(spam_mod, newdata = spam_test, type = "raw")
spam_pred_raw
spam_pred <- predict(spam_mod, newdata = spam_test, type = "class")
spam_pred_class
```

## Step 8: Model Verification
```{r modelVerification}
# Get Data from ClassExamples Directory
spam_test2 <- read_csv(file='EmailSpamFiltering/lingSpam.csv')
head(spam_test2)

spam_pred2 <- predict(spam_mod, newdata = spam_test2, type = "class")
spam_pred2
```

```{r seeResults}
# Print Confusion Matrix
spam_pred_table2 <- table(spam_test2$Label, spam_pred2)
spam_pred_table2

# Accuracy
sum(diag(spam_pred_table2)) / nrow(spam_test2)
```

## Step 9: Final Spam Identification Results

### Create confusion matrix of our results
```{r confusion}
# Print Confusion Matrix
spam_pred_table <- table(spam_test$Label, spam_pred)
spam_pred_table

# Accuracy
sum(diag(spam_pred_table)) / nrow(spam_test)
```


## References
https://www.kaggle.com/nitishabharathi/email-spam-dataset