---
title: "TS_MSFT_Stock"
author: "Isac Artzi"
date: "12/23/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Value of Microsoft stock between 2006 and 2018.

Data source: https://www.kaggle.com/szrlee/stock-time-series-20050101-to-20171231?select=MSFT_2006-01-01_to_2018-01-01.csv

### Acquire data
```{r set_data}
library(forecast)
stock_value <- read.csv("C:/GCU Academics/Junior Year/Second Semester/CST-425/CLC Shared GitHub Repository/Cloned Repository/ClassExamples/Topic-6/3-1-2022/MSFT_2006-01-01_to_2018-01-01.csv")
head(stock_value)
tail(stock_value)
```

### Create time series

```{r timeseries}


value.ts <- ts(stock_value$Close, start = c(2006,1,3),frequency = 365.25)

# Visualize the time series
plot(value.ts, xlab="Time", ylab="Stock value in $", ylim=c(0,100))
```

### Zoom in for additional insight

```{r zoom_in}
value.ts.zoomin <- ts(stock_value$Close, start = c(2011,1,1), end = c(2012,12,31),frequency = 365)

# Visualize the time series
plot(value.ts.zoomin, xlab="Time", ylab="Stock value in $", ylim=c(20,40))

```

### Verify seasonality

```{r seasonality}

fit <- tbats(value.ts)
seasonal <- !is.null(fit$seasonal)
seasonal

```

### Verify if there is a trend

After we concluded there is no seasonality, create a new time series object to verify presence of a trend

```{r trend}
# Plot Central Moving Average (CMA)
library(tsutils)

cma <- cmav(value.ts, outplot=1)

head(cma)

```
### Split the series into **trend**, **seasonality**, and **error**

```{r split_series}

decomposed <- decompose(value.ts, type="mult")
plot(decomposed)
stlRes <- stl(value.ts, s.window = "periodic")
lines(trendcycle(stlRes),col="red")

library(ggplot2)
autoplot(cbind(
	    Data=value.ts,
	    Seasonal=seasonal(stlRes),
  	  Trend=trendcycle(stlRes),
	    Remainder=remainder(stlRes)),
    facets=TRUE) +
  ylab("") + xlab("Day")

seasplot(value.ts)
```
### Prepare to forecast with simple exponential smoothing

Fit a simple exponential smoothing predictive model using the **Holt-Winters** model

```{r exp_smoothing}
# Plot the time series again
plot.ts(value.ts)

value.ts.forecast <- HoltWinters(value.ts, beta=FALSE, gamma=FALSE)
head(value.ts.forecast)
```
### Interpretation of Holt-Winters output

The estimated value of the alpha parameter is approximately 0.085. Since it is nearly zero, it follows that the forecasts are based on both recent and less recent observations:

```{r fitted}
# A sample of the first 10 fitted values
head(value.ts.forecast$fitted)

# Original time series and the forecast
plot(value.ts.forecast)

# Calculate the sum of square errors (SSE)
value.ts.forecast$SSE
```
### Make the forecast for a few days ahead using the exponential smoothing

```{r forecast}
# Forecast 200 days ahead with confidence intervals

value.ts.forecast2 <- forecast:::forecast.HoltWinters(value.ts.forecast, h=200)
head(value.ts.forecast2)

# The accuracy decreases as the time progresses into the future
plot(forecast(value.ts.forecast2))
```
### Prepare to forecast with ARIMA



```{r ARIMA}

# Forecast with auto.arima()

value.ts.arima <- arima(value.ts, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
value.ts.arima

library(lmtest)
coeftest(value.ts.arima)

predictions <- forecast:::forecast.Arima(value.ts.arima,h=200, level=c(99.5))
plot(forecast(predictions))
```
The blue line represents the forecast, while the gray area represeents 80% prediction intervals