---
title: "2-17-2022 - Team 1 Class Activity"
author: "Andrew Esch, Evan Lee, Collin Stratton"
date: "2-17-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load related packages

```{r libraries}
library(psych)
library(rela)   # contains Bartlett and KMO tests to verify assumptions
library(MASS)
library(parallel)

options(scipen=999)

# General filtering
library(dplyr)

# Low variance filter
library(creditmodel)

# Random forests
library(randomForest)
```

## Step 1: Identify a dataset (UCI, Kaggle, make one up, etc)
```{r prep_data}
# Andrew CD: C:/Users/andre/Documents/GitHub/CST-425/ClassExamples/Topic-5/2-17-2022/HighestGrossers.csv
data <- read.csv(file="C:/Users/andre/Documents/GitHub/CST-425/ClassExamples/Topic-5/2-17-2022/train.csv", header=TRUE,sep=",")
```

## Step 2: Explore the variables that make up the data
```{r exploreData}
head(data)
tail(data)
```

## Step 3: Formulate a predictive question

## Step 4: Build a model that predicts outcomes by combining a reduced number of essential variables (i.e. those which explain most of the variation)

## Dimensionality Reduction: Missing Values Ratio
```{r missingValues}
# Goal: Remove columns that have 1:7.5 ratio (at least 13.333%) of missing data to total data
#points <- vector("list", length(nor))

print(ncol(data))
for (i in 1:ncol(data)) {
  if (sum(is.na(data[,i])) / nrow(data) > 0.1333) {
    print(paste0("Remove column ", i)) # Specify column removed
    print(paste0("Ratio: ", sum(is.na(data[,i])) / nrow(data))) # Print ratio
    index <- i *-1
    data %>% select(index)
  }
}

#print(points)
#data %>% select(-1, -points)
print(ncol(data))

```

## Dimensionality Reduction: Low Variance Ratio
```{r lowVariance}
low_variance_filter(data, lvp = 0.97, only_NA = FALSE, note = FALSE, ex_cols = NULL)
```

## Dimensionality Reduction: Random Forests
```{r randomForests}
# TK
```


Formulate a predictive question about the data

Reduce the variables using three different methods and compare the results
Post your Markdown document that shows the reduced number of variables in the model, i.e. after you performed dimensionality reduction. Analysis refers to performing the reduction steps using methods described above

# References
https://www.tutorialspoint.com/r/r_random_forest.htm